---
title: "Happy Git and GitHub for the useR"
output: html_document
---

# Introduce yourself to Git

- substituting your name and the email associated with your GitHub account.

```
git config --global user.name "YangPing"
git config --global user.email "yangping1113@icloud.com"
git config --global --list
```

```{r}
library(usethis)
use_git_config(user.name = "YangPing", user.email = "yangping1113@icloud.com")
```

- You can check that Git understood what you typed by looking at the output of `git config --global --list`.


# Configure the default name for an initial branch

- set your default initial branch name to main like so, in the shell:
`git config --global init.defaultBranch main` or from R (the default for name is "main"):
```{r}
usethis::git_default_branch_configure()
```

---

# Basic shell commands

- `pwd`(print working directory) Shows directory or “folder” you are currently operating in. This is not necessarily the same as the R working directory you get from getwd().

- `ls`(list files). Shows the files in the current working directory. This is equivalent to looking at the files in your Finder/Explorer/File Manager. Use `ls -a` to also list hidden files, such as `.Rhistory` and `.git`.

- `cd` (change directory). Allows you to navigate through your directories by changing the shell’s working directory. You can navigate like so:
  - go to subdirectory foo of current working directory: cd foo
  - go to parent of current working directory: `cd ..`
  - go to your “home” directory: `cd ~` or simply `cd`
  - go to directory using absolute path, works regardless of your current working directory: `cd /home/my_username/Desktop`.

---

# A few Git commands:

- `git status` is the most used git command and informs you of your current branch, any changes or untracked files, and whether you are in sync with your remotes.
- `git remote -v` lists all remotes. Very useful for making sure git knows about your remote and that the remote address is correct.
- `git remote add origin GITHUB_URL` adds the remote GITHUB_URL with nickname origin.
- `git remote set-url origin GITHUB_URL` changes the remote url of origin to GITHUB_URL. This way you can fix typos in the remote url.

---

# Generate token

Go to https://github.com/settings/tokens and click “Generate token”.
Or, from R, do:
```{r}
usethis::create_github_token()
```
Look over the scopes; I highly recommend selecting “repo”, “user”, and “workflow”.

You could even get out ahead of this and store the PAT explicitly right now. In R, call gitcreds::gitcreds_set() to get a prompt where you can paste your PAT:
```{r}
gitcreds::gitcreds_set()
```

---

# PAT
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

Changing a specific remote from HTTPS to SSH (and back again) is a straightforward operation with git remote set-url REMOTE_NAME DESIRED_URL:
```
~/rrr/happy-git-with-r % git remote -v
origin  https://github.com/jennybc/happy-git-with-r.git (fetch)
origin  https://github.com/jennybc/happy-git-with-r.git (push)

~/rrr/happy-git-with-r % git remote set-url origin git@github.com:jennybc/happy-git-with-r.git

~/rrr/happy-git-with-r % git remote -v
origin  git@github.com:jennybc/happy-git-with-r.git (fetch)
origin  git@github.com:jennybc/happy-git-with-r.git (push)

~/rrr/happy-git-with-r % git remote set-url origin https://github.com/yangping2022/happy-git.git
```
We can do the same from R using functions in usethis:
```
usethis::git_remotes()
#> $origin
#> [1] "https://github.com/yangping2022/happy-git.git"

usethis::use_git_remote(
  "origin",
  "git@github.com:jennybc/happy-git-with-r.git",
  overwrite = TRUE
)

usethis::git_remotes()
#> $origin
#> [1] "git@github.com:jennybc/happy-git-with-r.git"

usethis::use_git_remote(
  "origin",
  "https://github.com/yangping2022/happy-git.git",
  overwrite = TRUE
)
```

---

# Clone the repo to your local computer

Take charge of – or at least notice! – what directory you’re in. `pwd` displays the working directory. `cd` is the command to change directory. Personally, I would do this sort of thing in ~/tmp.

Clone myrepo from GitHub to your computer. Use the URL we just copied from GitHub. This URL should have **your GitHub username** and the name of your **practice repo**. If your shell (Appendix A) cooperates, you should be able to paste the whole https://.... bit that we copied above. But some shells are not (immediately) clipboard aware. In that sad case, you must type it. **Accurately**.
```
git clone https://github.com/YOUR-USERNAME/YOUR-REPOSITORY.git
```

Make this new repo your working directory, list its files, display the README, and get some information on its connection to GitHub:
```
cd myrepo
ls
head README.md
git remote show origin
```


`pwd` displays the working directory. `cd` is the command to change directory. Personally, I would do this sort of thing in `~/tmp`.

```
~/tmp % cd myrepo

~/tmp/myrepo % ls
README.md

~/tmp/myrepo % head README.md 
# myrepo
checking stuff for Happy Git

~/tmp/myrepo % git remote show origin
* remote origin
  Fetch URL: https://github.com/yangping2022/happy-git.git
  Push  URL: https://github.com/yangping2022/happy-git.git
  HEAD branch: main
  Remote branch:
    main tracked
  Local branch configured for 'git pull':
    main merges with remote main
  Local ref configured for 'git push':
    main pushes to main (up to date)
```


#  Make a local change, commit, and push

```
~/tmp/myrepo % echo "A line I wrote on my local computer" >> README.md

~/tmp/myrepo % git status
```

If you already stored your PAT with `gitcreds::gitcreds_set()`, it should be discovered automatically and you will not see a credential challenge.
```
git add README.md
git commit -m "A commit from my local computer"
git push
```

---

# Clean up

Here’s how to do that in the shell, if current working directory is myrepo:
```
cd ..
rm -rf myrepo/
```

---

# Connect RStudio to Git and GitHub

Clone the test GitHub repository to your computer via RStudio
In RStudio, start a new Project:

- *File > New Project > Version Control > Git.* In “Repository URL”, paste the URL of your new GitHub repository. It will be something like this `https://github.com/yangping2022/happy-git.`

  - Do you NOT see an option to get the Project from Version Control? Restart RStudio and try again. Still no luck? Go to chapter 13 for tips on how to help RStudio find Git.

- Accept the default project directory name, e.g. myrepo, which coincides with the GitHub repo name.

- Take charge of – or at least notice! – where the Project will be saved locally. A common rookie mistake is to have no idea where you are saving files or what your working directory is. Pay attention. Be intentional. Personally, I would do this in `~/tmp`.

- I suggest you check “Open in new session”, as that’s what you’ll usually do in real life.

- Click “Create Project”.

You should find yourself in a new local RStudio Project that represents your test repo on GitHub. This should download the README.md file from GitHub. Look in RStudio’s file browser pane for the `README.md` file.

---

# Make local changes, save, commit.  

From RStudio, modify the README.md file, e.g., by adding the line “This is a line from RStudio”. Save your changes.

Commit these changes to your local repo. How?
From RStudio:
- Click the “Git” tab in upper right pane.
- Check “Staged” box for README.md.
- If you’re not already in the Git pop-up, click “Commit”.
- Type a message in “Commit message”, such as “Commit from RStudio”.
- Click “Commit”.

# Push your local changes online to GitHub
Click the green “Push” button to send your local changes to GitHub.

---

# Optional: peek under the hood

- `git remote -v` or `git remote --verbose` shows the remotes you have setup. Here’s how that looks for someone using HTTPS with GitHub and calling it origin:
```
~/tmp/myrepo % git remote -v
origin  https://github.com/yangping2022/happy-git.git (fetch)
origin  https://github.com/yangping2022/happy-git.git (push)
```
- `git branch -vv` prints info about the current branch (-vv for “very verbose”, I guess). In particular, we can see that local main is tracking the main branch on origin, a.k.a. origin/main.
```
~/tmp/happy-git % git branch -vv
* main b6f06af [origin/main] pull from RStudio
```
- Finally, git remote show origin gives yet another view on useful remote and branch information:
```
~/tmp/happy-git % git remote show origin
```
git clone, which RStudio did for us, sets all of this up automatically. This is why “GitHub first, then RStudio” is the preferred way to start projects early in your Git/GitHub life.

# From RStudio, modify the README.md file, e.g., by adding the line “This is a line from RStudio”. 
Save your changes.
Commit these changes to your local repo. How?
- Click the “Git” tab in upper right pane
- Check “Staged” box for any files whose existence or modifications you want to commit.
- To see more detail on what’s changed in file since the last commit, click on “Diff” for a Git pop-up
- If you’re not already in the Git pop-up, click “Commit”
- Type a message in “Commit message”, such as “Commit from RStudio”.
- Click “Commit”

# Confirm the local change propagated to the GitHub remote

- Go back to the browser. I assume we’re still viewing your new GitHub repo.
- Refresh.
- You should see the new “This is a line from RStudio” in the README.
- If you click on “commits,” you should see one with the message “Commit from RStudio”.

# Make a change on GitHub
Click on README.md in the file listing on GitHub.

In the upper right corner, click on the pencil for “Edit this file”.

Add a line to this file, such as “Line added from GitHub.”

Edit the commit message in “Commit changes” or accept the default.

Click the big green button “Commit changes.”

# Note that in general (and especially in future when collaborating with other developers) you will usually need to pull changes from the remote (GitHub) before pushing the local changes you have made. For this reason, it’s a good idea to try and get into the habit of pulling before you attempt to push.

---

# Existing project, GitHub last
## Make or verify an RStudio Project
If the project is not already an RStudio Project, make it so:
- Within RStudio you can do: File > New Project > Existing Directory and, if you wish, “Open in new session”.
- Alternatively, from R, call `usethis::create_project("path/to/your/project")`, substituting the path to your existing project directory.
- If your project is already an RStudio Project, launch it.

## Make or verify a Git repo
You should be in RStudio now, in your project.

Is it already a Git repository? The presence of the Git pane should tip you off. If yes, you’re done.

If not, you have several options:

- In the R Console, call `usethis::use_git()`.
- In RStudio, go to *Tools > Project Options … > Git/SVN*. Under “Version control system”, select “Git”. Confirm New Git Repository? Yes!
- In the shell, with working directory set to the project’s directory, do `git init`.
If you used usethis or RStudio to initialize the Git repo, the Project should re-launch in RStudio. Do that yourself if you did git init. RStudio should now have a Git pane.

# stage and commit
If your local project was already a Git repo and was up-to-date, move on. Otherwise, you probably need to stage and commit.
- Click the “Git” tab in upper right pane
- Check “Staged” box for all files you want to commit.
- Default: stage everything
- When to do otherwise: this will all go to GitHub. So consider if that is appropriate for each file. You can absolutely keep a file locally, without committing it to the Git repo and sending to GitHub. Just let it sit there in your Git pane, without being staged. No harm will be done. If this is a long-term situation, list the file in .gitignore.
- If you’re not already in the Git pop-up, click “Commit”
- Type a message in “Commit message”.
- Click “Commit”

---

# Make and connect a GitHub repo with usethis
In your project, in the R Console, call:
```
usethis::use_github()
```

---

# In the shell
In a shell, do this, substituting your URL:
```
git remote add origin  https://github.com/yangping2022/happy-git.git
```
Push and cement the tracking relationship between your local main branch and main on GitHub (or whatever your default branch is named):
```
git push --set-upstream origin main
```

---

# Test drive R Markdown
We’ll practice with RStudio’s boilerplate R Markdown document.

Do this: *File > New File > R Markdown …*

- Give it an informative title. 
- Accept the default Author or edit if you wish.
- Accept the default output format of HTML.
- Click OK.
Save this document to a reasonable filename and location. The filename should end in `.Rmd` or `.rmd`. Save in the top-level of this RStudio project and Git repository, that is also current working directory. 
Click on **“Knit HTML” or do File > Knit Document**. RStudio should display a preview of the resulting HTML. Also look at the file browser. You should see the original R Markdown document, i.e. `foo.Rmd` AND the resulting HTML `foo.html`.

# Output format

The magical process that turns your R Markdown to HTML is like so:
```
foo.Rmd --> foo.md --> foo.html
```
switch the output format to github_document. This means rendering look like this:
```
foo.Rmd --> foo.md
```
- You can make some YAML changes via the RStudio IDE: click on the “gear” in the top bar of the source editor, near the “Knit HTML” button. Select **“Output options”** and go to the Advanced tab and check **“Keep markdown source file”**. Your YAML should now look more like this:
```
---
title: "Something fascinating"
author: "Jenny Bryan"
date: "2022-03-25"
output:
  html_document:
    keep_md: true
---
```
- You should have gained the line keep_md: true. You can also simply edit the file yourself to achieve this. The IDE only exposes a small fraction of what’s possible to configure in the YAML.

- In fact, a hand-edit is necessary if you want to keep only markdown and get GitHub-flavored markdown. In that case, make your YAML look like this:
```
---
title: "Something fascinating"
author: "Jenny Bryan"
date: "2022-03-25"
output: github_document
---
```
Save!
- Render via “Knit HTML” button.
Now revisit the file browser. In addition to `foo.Rmd`, you should now see `foo.md`. If there are R chunks that make figures, the usage of markdown output formats will also cause those figure files to be left behind in a sensibly named sub-directory, such as foo_files.

# Put your stamp on it

Insert an empty R chunk, via the “Chunk” menu in upper right of source editor or with the corresponding keyboard shortcut.
```{r, eval=TRUE}
## insert your brilliant WORKING code here
```
If the code doesn’t work with you babysitting it, I can guarantee you it will fail, in a more spectacular and cryptic way, when run at arms-length via “Knit HTML” or `rmarkdown::render()`.

# Publish your report

Use `output: github_document` or, if output is html_document, add `keep_md: true`.
**In both cases, it is critical to also commit and push everything inside** `foo_files`

# HTML on GitHub

HTML files, such as `foo.html`, are not immediately useful on GitHub
But there are ways to get a preview: such as
https://htmlpreview.github.io/

---

# If you can’t figure out what’s wrong by reading the error messages
- Insert this chunk near the top of your .Rmd document:
```
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```
```
- If it’s undesirable to globally accept errors, you can still specify error = TRUE for a specific chunk like so:
```
```{r wing-and-a-prayer, error = TRUE}
## your sketchy code goes here ;)
```
```
- Adapt the “git bisect” strategy:
Put `knitr::knit_exit()` somewhere early in your `.Rmd` document, either in inline R code or in a chunk. Keep moving it earlier until things work. Now move it down in the document. Eventually you’ll be able to narrow down the location of your broken code well enough to find the line(s) and fix it.
- Check your working directory
  - `getwd()` will display working directory at run time.
  - `list.files()` will list the files in working directory. Is the file you want even there?
  
#  Render an R script
## Transform the R markdown to R:
- Anything that’s not R code? Like the YAML and the prose? Protect it with roxygen-style comments: start each line with #'
- Anything that is R code? Let it exist “as is” as top-level code. That means you’ll need to change the syntax of R chunk headers like so:
Before: ````{r setup, include = FALSE}`
After: `#+ r setup, include = FALSE`
  
```{r}
dotchart(VADeaths, main = "Death Rates in Virginia - 1940")
```
  
  